Music Perception: Analysis

### 2021-02-08 analysis/analysis.Rmd
move previous analysis to pilot folder
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception: Pilot Accuracy Analysis"
author: "Atsuko Tominaga"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

- 10 participants (experimental error = 1) >> 9 participants were included for analysis.

# Experiment

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-1](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-1)).
- 48 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 96 stimuli). The half of them were selected from the teaching condition and the other half were from the performing condition.
  + 24 for teaching-articulation performances
  + 24 for performing-articulation performances
  + 24 for teaching-dynamics performances
  + 24 for performing-dynamics performances

## Design

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two expressions separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomised across participants.

## Procedure
1. Musicians (> 6yo experience) were asked to listen to a number of recordings and judge whether each performance was produced for teaching purposes or not.

Actual instruction:
```
Each performance was produced in order to either1) teach the musical expressive technique (as a teacher) 2) perform their best (as a performer).

You will be asked to judge whether each performer had the intention to teach or not by pressing the 'Yes' <Left> or 'No' <Right> key.
```

2. Participants pressed "YES" if they think the recording was produced for teaching (as a teacher) whereas they pressed "NO" if they thought the recording was produced for performing (as a performer).

3. Correct answers would be 1) participants pressed "YES" when they listened to teaching performances or 2) participants pressed "NO" when they listened to performing performances.

```{r packages, include = FALSE}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("afex")) {install.packages("afex"); require("afex")}
```

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)
```

# Results
## Individual data
```{r summary, echo = FALSE}
# remove irrelevant strings
data <- all_data[expMode == "experiment"]
data$MidFile <- gsub("./mid/", "", data$midFile)
data$MidFile <- gsub(".mid", "", data$MidFile)

# name conditions
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", midFile)]$Condition <- "teaching"
data[grepl("_p_", midFile)]$Condition <- "performing"
data[grepl("a_", midFile)]$Skill <- "articulation"
data[grepl("d_", midFile)]$Skill <- "dynamics"

# check correct or incorrect
data$Correct <- 0
data[Condition == "teaching" & rating == "Yes"]$Correct <- 1
data[Condition == "performing" & rating == "No"]$Correct <- 1

all <- data[expMode == "experiment", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
all[order(subjectNumber, Condition, Skill)]
```

## 1. Summary
### All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/96)*100, SD_Percent = (sd(Sum)/96)*100)]
subset
```

```{r plot, echo = FALSE}
ggboxplot(ind_data, x = "N", y = "Percent", add = "dotplot", xlab = FALSE) + ylab("Correct %") + theme(axis.text.x = element_blank())
```

### One-sample *t*-test
```{r t-test, echo = FALSE}
# normality check
shapiro.test(ind_data$Percent)

t.test(ind_data$Percent, mu = 50, alternative = "two.sided")
```

### Articulation vs. Dynamics

```{r plot2, echo = FALSE}
ind_data_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Skill)]
ind_data_ad <- ind_data_ad[order(subjectNumber, Skill)]

subset_ad <- ind_data_ad[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(Skill)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad$subjectNumber, articulation = ind_data_ad[Skill == "articulation"]$Percent, dynamics = ind_data_ad[Skill == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## One-sample *t*-test
### Articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "articulation"]$Percent)

t.test(ind_data_ad[Skill == "articulation"]$Percent, mu = 50, alternative = "two.sided")
```
- Not significantly different from chance (p > .2).

### Dynamics
```{r t-test-d, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent)

t.test(ind_data_ad[Skill == "dynamics"]$Percent, mu = 50, alternative = "two.sided")
```

## Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent-ind_data_ad[Skill == "articulation"]$Percent)

t.test(ind_data_ad[Skill == "articulation"]$Percent, ind_data_ad[Skill == "dynamics"]$Percent, paired = TRUE)
```

## 2. Teaching stimuli
### All
```{r item1, echo = FALSE}
# individual
ind_data2 <- data[subjectNumber != 9 & Condition == "teaching", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill, MidFile)]
ind_data2 <- ind_data2[order(subjectNumber, Condition, Skill, MidFile)]

# all
subset2 <- ind_data2[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
subset2[order(-Mean_Percent)]
```

### Articulation
```{r item1-a, echo = FALSE}
subset2_a <- subset2[Skill == "articulation"]
subset2_a$MidFile <- gsub("a_t_", "", subset2_a$MidFile)
subset2_a[order(-Mean_Percent)]

bar_a <- ggplot(subset2_a, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_a
```

### Dynamics
```{r item1-d, echo = FALSE}
subset2_d <- subset2[Skill == "dynamics"]
subset2_d$MidFile <- gsub("d_t_", "", subset2_d$MidFile)
subset2_d[order(-Mean_Percent)]

bar_d <- ggplot(subset2_d, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_d
```

## 3. Performing stimuli
### All
```{r item2, echo = FALSE}
# individual
ind_data3 <- data[subjectNumber != 9 & Condition == "performing", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill, MidFile)]
ind_data3 <- ind_data3[order(subjectNumber, Condition, Skill, MidFile)]

# all
subset3 <- ind_data3[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
subset3[order(-Mean_Percent)]
```

### Articulation
```{r item2-a, echo = FALSE}
subset3_a <- subset3[Skill == "articulation"]
subset3_a$MidFile <- gsub("a_p_", "", subset3_a$MidFile)
subset3_a[order(-Mean_Percent)]

bar2_a <- ggplot(subset3_a, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_a
```

### Dynamics
```{r item2-d, echo = FALSE}
subset3_d <- subset3[Skill == "dynamics"]
subset3_d$MidFile <- gsub("d_p_", "", subset3_d$MidFile)
subset3_d[order(-Mean_Percent)]

bar2_d <- ggplot(subset3_d, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_d
```

```{r export, include = FALSE}
order_data <- rbind(subset2, subset3)

# export
fwrite(order_data, file = "order_data.txt")
```

>>>>

---
title: "Music Perception: Analysis"
output: html_notebook
---
